---
title: "Submetering"
author: "Murilo Miranda"
date: "3/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE}
library(RMySQL)
library(tidyverse)
library(zoo)
```

## Obtain the data
```{r}
con <- dbConnect(MySQL(), user='deepAnalytics', password='Sqltask1234!', dbname='dataanalytics2018', host='data-analytics-2018.cbrosir2cswx.us-east-1.rds.amazonaws.com')

# list the tables contained in the dataanalytics2018 database
dbListTables(con)

# list the attributes in the yr_2006 table
dbListFields(con, "yr_2006")

# query the attributes
query_data <- paste0("SELECT Date, Time, Sub_metering_1, Sub_metering_2, Sub_metering_3 FROM")

# fetch tables from 2006 till 2010
#dfList <- lapply(paste0("yr_", 2006:2010), 
#                 function(t) 
#                   dbGetQuery(con, paste(query_data, t)))

yr_2006 <- dbGetQuery(con, paste(query_data, "yr_2006"))
yr_2007 <- dbGetQuery(con, paste(query_data, "yr_2007"))
yr_2008 <- dbGetQuery(con, paste(query_data, "yr_2008"))
yr_2009 <- dbGetQuery(con, paste(query_data, "yr_2009"))
yr_2010 <- dbGetQuery(con, paste(query_data, "yr_2010"))
rm(query_data)
```

#### Check each table
```{r}
investigateTable <- function(data_table){
  cat("Structure:\n")
  cat(str(data_table))
  cat("\nSummary:\n")
  print(summary(data_table))
  cat("\nHead:\n")
  print(head(data_table))
  cat("\nTail:\n")
  print(tail(data_table))
}

# yr_2006:yr_2010
investigateTable(yr_2006) # from 16-12 till 31-12-2006
investigateTable(yr_2007) # full year
investigateTable(yr_2008) # full year
investigateTable(yr_2009) # full year
investigateTable(yr_2010) # full year
```

We combine all tables into one, except yr_2006 table because it only has few data from 16 until 31 December.

```{r}
submeter_data <- bind_rows(yr_2007, yr_2008, yr_2009, yr_2010)
rm(yr_2006, yr_2007, yr_2008, yr_2009, yr_2010)

investigateTable(submeter_data)
```

Date and Time attributes are character data type. We convert these attributes into DateTime type, but first we combine the two attributes in one and then convert.

```{r}
#combine the two attributes and convert the data type
submeter_data <- submeter_data %>% 
  mutate(datetime = as.POSIXct(paste(Date, Time, sep = "T"), tz = "Europe/Paris")) %>% 
  select(datetime, starts_with("Sub_"))

str(submeter_data)
```

```{r}
# verify missing values
head(submeter_data[which(is.na(submeter_data$DateTime)),])
```

There are four blocks of missing values: (1) 119637 - 119696, (2) 649945 - 650004, (3) 1173908 - 1173967 and (4) 1688695 - 1688754. They correspond to Daylight savings time (DST): (1) 2007-03-25 from 2:00:00 until 2:59:00, (2) 2008-03-30 from 2:00:00 until 2:59:00, (3) 
2009-03-29 from 2:00:00 until 2:59:00 and (4) 2010-03-28 from 2:00:00 until 2:59:00. These missing values were omit/remove.

```{r}
# check the previous and next times
submeter_data[c(119636:119637, 119696:119697), ]
```

We remove the missing DateTime values using the code below:

```{r}
submeter_data <- submeter_data %>% filter(!is.na(DateTime))
```

Now we verify if there is data which was not added into the database. For this reason, we make a vector (called *date_Time*) from `min(submeter_data$DateTime)` to ```max(submeter_data$DateTime)` by minute. After that, we merge the *submeter_data* database with *date_Time* vector. If there is any missing value, it will be imputed by interpolation.

```{r}
date_Time <- seq(min(submeter_data$DateTime), max(submeter_data$DateTime), "min")

##
submeter_data <- submeter_data %>% 
  full_join(as.data.frame(date_Time), by = c("DateTime" = "date_Time"))

submeter_data %>% ggplot(aes(DateTime, Sub_metering_1)) + geom_line()
```

